---
title: "Peer assignment 1-Reproducibile Research"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

First I load the data and process it.

```{r }
data <- read.csv("activity.csv")
library(dplyr)
library(ggplot2)
data %>% group_by(date) %>% summarize(stepsperday = sum(steps))
```

## What is the mean total number of steps taken per day?

What is mean total number of steps taken per day? For this part, ignoring the missing value, I calculate the total number of steps taken per day and then put it into a histogram. After this exploratory phase, I calculated the mean and median of the total number of steps taken per day.

```{r}
data %>% group_by(date) %>% summarize(stepsperday = sum(steps)) %>% ggplot(aes(x = stepsperday)) + geom_histogram(bins = 20) + ggtitle("Total number of steps per day") + xlab("Steps per day")

mean(data$steps, na.rm= TRUE)
median(data$steps, na.rm = TRUE)

m1 <- with(data, tapply(steps, date, mean, na.rm = TRUE))
m2 <- with(data, tapply(steps, date, median, na.rm = TRUE))
d1 <- data.frame(date = names(m1), mean = m1)
d2 <- data.frame(date= names(m2), median = m2)
par(mfcol = c(1, 2))
with(d1, plot(date, mean, main = "Mean steps per day", xlab = "Time", ylab = "Mean"))
with(d2, plot(date, median, main = "Median steps per day", xlab = "Time", ylab = "Median"))
```

## What is the average daily activity pattern?

I create a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days. I calculated which 5-min interval contains the max number of steps.

```{r}
averageAct <- aggregate(data$steps, by = list(data$interval), FUN = mean, na.rm = TRUE)
names(averageAct) <- c("interval", "average")

plot(averageAct$interval, averageAct$average, type = "l", col = "steelblue", lwd = 2, xlab = "Intervals", ylab = "Average n of steps")

max <- max(data$steps, na.rm = TRUE)
max
```

## Imputing missing values

In this part I've imputed the missing values, assigning to them the values of the mean per interval. After imputing, I've created a histogram of the total number of steps taken each day.

```{r}
NAs <- is.na(data$steps)
table(NAs)

data2 <- averageAct$average[match(data$interval, averageAct$interval)]
imputed <- transform(data, steps = ifelse(is.na(data$steps), yes = data2, no = data$steps))
total_imputed <- aggregate(steps ~ date, imputed, sum)

hist(total_imputed$steps, xlab = "Total steps per day", main = "Total number of steps taken each day", breaks = 10, col = "steelblue", ylim = c(0, 30))

mean_NoNA <- mean(total_imputed$steps)
mean_NoNA

median_NoNA <- median(total_imputed$steps)
median_NoNA
```

## Are there differences in activity patterns between weekdays and weekend?

I've created a plot containing a time series of the 5-min interval and the average number of steps taken, divided by weekdays and weekends

```{r}
data$date <- as.Date(strptime(data$date, format = "%Y-%m-%d"))
data$datetype <- sapply(data$date, function (x){
    if (weekdays(x) == "sabato" | weekdays(x) == "domenica"){
        y <- "Weekend"} else {
            y <- "Weekday"}
    y
})


data3 <- aggregate(steps ~ interval + datetype, data, mean, na.rm = TRUE)
ggplot(data3, aes(x=interval, y=steps, color=datetype)) + geom_line() + facet_wrap(~datetype, ncol = 1, nrow= 2) + labs(title = "Average daily steps by type of date", x= "Interval", y= "Average number of steps")
```